---
title: "Statistically Modelling the Impact of Macroeconomic factors on the Economic growth in the United States using Linear Regression"
author: "Bridget _Chukwu"
date: "June 20, 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Uploading the data and cleaning the data 

```{r}
rm(list=ls())
getwd()
setwd("C:\\Users\\bridget.chukwu\\Documents\\Summer Course STAT\\STAT 711\\Final Project")
library(readxl)
data <- read_excel("Data for final Project.xlsx")
meaningful_columns <- c("Year", "GDP per capita growth (annual %)", "FDI(% of GDP)", "Real interest rate (%)", "Tax revenue (% of GDP)")
cleaned_data <- data[meaningful_columns]
data<-cleaned_data 
data<-na.omit(data)
dim(data)
head(data)
names(data)
data
nrow(data)
ncol(data)
FDI<-data$"FDI(% of GDP)"
GDP<-data$"GDP per capita growth (annual %)"
TAX<-data$"Tax revenue (% of GDP)"
INT<-data$"Real interest rate (%)" 
Year<-data$Year
```
 
 ##   Pre-estimation Test and Summary Statistics
```{r}
summary(data)
```

#### Standard Deviation 
```{r}

sd(GDP)
sd(FDI)
sd(INT)
sd(TAX)

```
 
#### Ploting line graph for the variables 

```{r}
par(mfrow = c (2, 3))
LGDP <- log(data$`FDI(% of GDP)` + 1)
plot(Year, GDP, main="Time Trend for GDP)", type = "l", col = "Red", ylim = c(-10,10), xlab = "Year", ylab = "GDP" )
plot(Year, LGDP, main="Time Trend for LOG GDP", type = "l", col = "Red", ylim = c(-10,10), xlab = "Year", ylab = "LGDP" )
names(data)

plot(Year, FDI, main="Time Trend for Foreign Direct investment", type = "l", col = "Red", ylim = c(-10,10), xlab = "Year", ylab = "FDI" )
plot(Year, INT, main="Time Trend for real Interest Rate", type = "l", col = "Red", ylim = c(-10,20), xlab = "Year", ylab = "INT" )
plot(Year, TAX, main="Time Trend for Tax Revenue", type = "l", col = "Red", ylim = c(-10,20), xlab = "Year", ylab = "TAX" )
```

#### Plotting Density Plot plot
```{r}
par(mfrow=c(2,2))
plot(density(LGDP))
plot(density(FDI))
plot(density(TAX))
plot(density(INT))
```

#### Plotting the Q-Q plots
```{r}
par(mfrow=c(2,2))
qqnorm(LGDP)
qqline(LGDP)
qqnorm(FDI)
qqline(FDI)
qqnorm(TAX)
qqline(TAX)
qqnorm(INT)
qqline(INT)

```

```{r}
# Scatter plots for the data variables 
par(mfrow=c(2,3))
plot(LGDP,FDI)
plot(LGDP,TAX)
plot(LGDP,INT)

```

# Scatter plots for the data variables 
```{r}
par(mfrow=c(2,3))
plot(LGDP,FDI)
plot(LGDP,TAX)
plot(LGDP,INT)
```

## Correlation between the variables
```{r}
cor(LGDP,FDI)
cor(LGDP,TAX)
cor(LGDP,INT)
```

### Linear Regression without interactions
```{r}

result<-lm(LGDP~FDI+TAX+INT,data = data);result
summary(result)
```

## # Best linear model without interaction terms using backward reduction
```{r}

reduce<-step(result,direction = "backward")
summary(reduce)
```

# Best linear model without interaction terms using forward selection
```{r}
min<-lm(GDP~1)
forward<-step(min,direction = "forward",scope = LGDP ~ FDI + TAX + INT)
summary(forward)
```

## Linear Regression Result with two interactions terms
```{r}

result2<-lm(LGDP~(FDI+TAX+INT)^2,data = data);result2
summary(result2)

```

## Best linear model with two interaction terms using backward reduction
```{r}
reduce2<-step(result2,direction = "backward")
summary(reduce2)
```

